---
title: "p8105_hw2_jn2761"
author: "jiaying Ning"
date: "9/24/2020"
output: html_document
---

```{r}
library(tidyverse)
library(readxl)
library(janitor)

```

##problem

Read the Mr. Trashwheel dataset.
```{r}
#import data and specify range of column
Trashweel_df=
   read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
            sheet = "Mr. Trash Wheel",
            range = cell_cols("A:N"))%>% 
              janitor::clean_names() %>% #clean variable name
              drop_na(dumpster) %>% 
  #only drop na for that column,omit rows that do not include dumpster-specific data
              mutate(
                sports_balls=round(sports_balls),
                sports_balls=as.integer(sports_balls)
                #round the number of sports balls to the nearest integer and converts the result to an integer variable (using as.integer)
              )
```

Read Precipitation Date

```{r}
Precip_2018=
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
            sheet = "2018 Precipitation",
            skip = 1
            )%>%
            janitor::clean_names() %>%
  drop_na(month) %>%
            mutate(year = 2018) %>%
            relocate(year)
            
Precip_2017=
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
            sheet = "2017 Precipitation",
                  skip = 1
            )%>%
            janitor::clean_names() %>%
  drop_na(month) %>%
            mutate(year = 2017) %>%
              relocate(year)
            
#Looking at excel sheet, we should skip the first show
```


## Problem 2
This problem focuses on NYC Transit data; in particular, this CSV file contains information related to each entrance and exit for each subway station in NYC.

Read and clean the data; retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. Convert the entry variable from character (YES vs NO) to a logical variable (the ifelse or recode function may be useful).

```{r import data}
NYC_Transitdf = 
  read.csv(
    "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv"
  ) %>%
            janitor::clean_names() %>%
  select(line:entry,ada,vending) %>%
   mutate( entry = ifelse( entry =="NO", FALSE, TRUE)) %>%
  mutate(
    route8 = as.character(route8),
    route9 = as.character(route9),
    route10 = as.character(route10),
    route11 = as.character(route11)
    )

```



The current dataset contain variable `r names(NYC_Transitdf)`. So far we clean the name of each variable and kept only the variable of interest which are `r names(NYC_Transitdf)`. We also convert the entry variable from character to logical. After importing and cleaning, we now have a dataset that have dimensions of `r nrow(NYC_Transitdf)`X`r ncol(NYC_Transitdf)`. These data is not tidy, in this case the routes served spread across 10 columns.


Answer the following questions using these data:


```{r tidy}

#create data frame that tidy up the route information
NYC_Transit_tidydf = 
    pivot_longer(
  NYC_Transitdf,
    route1:route11, 
    names_to = "route",
    values_to = "train")%>%
  unite(station_name_unique, c(line, station_name),remove=FALSE)

```


```{r}
#create data frame that contain distinct station name for each row
NYC_Transit_distinct = distinct(NYC_Transit_tidydf, station_name_unique,.keep_all =  TRUE)
#How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St A/B/C/D; 125st 1; 125st 4/5); the distinct function may be useful here.
nrow(distinct(NYC_Transit_distinct,station_name_unique))
#How many stations are ADA compliant?
nrow(subset(NYC_Transit_distinct, ada=="TRUE"))
#What proportion of station entrances / exits without vending allow entrance?
prop.table(table(NYC_Transit_distinct$vending))
#Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?
nrow(subset(NYC_Transit_distinct, train=="A"))
nrow(subset(NYC_Transit_distinct, train=="A"& ada=="TRUE"))

```

**Answer**: There are `r nrow(distinct(NYC_Transit_distinct,station_name_unique))` distinct stations. and `r nrow(subset(NYC_Transit_distinct, ada=="TRUE"))`  stations are ADA compliant. 0.01935484 of the station/ exits without vending allow entrance.There are `r nrow(subset(NYC_Transit_distinct, train=="A"))`
 distinct stations serve the A train? Of the stations that serve the A train, and `r nrow(subset(NYC_Transit_distinct, train=="A"& ada=="TRUE"))` of them are ADA compliant?



## Problem 3

This problem uses the FiveThirtyEight data; these data were gathered to create the interactive graphic on this page. In particular, we’ll use the data in pols-month.csv, unemployment.csv, and snp.csv. Our goal is to merge these into a single data frame using year and month as keys across datasets.

First, clean the data in pols-month.csv. separate() is used to break up the variable mon into integer variables year, month, and day; replace month number with month name; create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable.

```{r}
pols_month = 
  read.csv(
    "./data/pols-month.csv"
  ) %>% 
  separate(mon, c("year","month","day"), sep = "([-])") %>% 
  select(-day,-prez_dem,-prez_gop)%>%
  mutate(
    year = as.integer(year),
    month = as.integer(month),
    month = month.abb[month],
gop = gov_gop + sen_gop +rep_gop,
dem = gov_dem + sen_dem +rep_dem)

pols_month_data = 
  pivot_longer(
    pols_month, 
    gov_gop:rep_dem,
    names_to = "president", 
    values_to = "pols")


```


Second, clean the data in snp.csv using a similar process to the above. For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.

```{r}
snp = 
  read.csv(
    "./data/snp.csv"
  ) %>% 
  separate(date, c("day","month","year"), sep = "([/])")%>% 
  select(-day)%>%
  mutate(
    year = as.integer(year),
    month = as.integer(month),
    month = month.abb[month])%>% 
  relocate(year)

```


Third, tidy the unemployment data so that it can be merged with the previous datasets. This process will involve switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.

```{r}
unemployment = 
  read.csv(
    "./data/unemployment.csv"
  )   

unemployment_data = 
  pivot_longer(
    unemployment, 
    Jan:Dec,
    names_to = "month", 
    values_to = "unemployment_rate")%>%  janitor::clean_names()



```


Join the datasets by merging snp into pols, and merging unemployment into the result.

```{r}
pols <- merge(pols_month_data, snp, by.x = c('year', 'month'), by.y = c('year', 'month'))
pols <- merge(pols, unemployment_data, by.x = c('year', 'month'), by.y = c('year', 'month'))

summary(pols)

```




The dataset *pols_month*  contains in total of 822 observations and 9 variable, it recorded the number of democratic and republican at a given time.

The dataset *snp*  contains in total of 787 observations and 2 variable, it represent measure of stock market(closing values of the S&P stock index) at a given time.

the dataset *unemployment* contains 68 obersvation and 13 variable, it recorded the percentage of unemployment rate at a given time.

IN the resulting dataset, I have 4722 observation and 8 variables, years range from 1950 to 2015, some of the key variables are number of republican(gop), number of democrate (dem), stock market measure (close), and umemployment_rate.






